<P>

<BR>

<CENTER>

<FONT SIZE = 6, COLOR="#000066"> <B> ALY 6015 (Intermediate Analytics)</Br>

Team Project Phase 2: Preliminary Analysis Report </Br> Group 5 </Br></FONT>

Team Members: Sankalp Susil Kumar Biswal, Sanchi Gupta, Sharanya Badrinarayanan, Ratnesh Mishra, Rhea John Thoppil </Br></FONT>

<FONT SIZE = 4, COLOR="#660000"> Instructor: Vladimir Shapiro, Northeastern University </FONT>

</CENTER>

</B>

# I. Introduction

The exploration of the FIFA 2022 dataset provides a statistical lens through which one can gauge and predict football player performance. By leveraging key performance indicators such as pace, shooting, and passing, we aim to employ regression analysis to not only predict overall performance but also delve into position-specific skills. The research questions poised are designed to uncover the depths of player abilities and predict market value---elements crucial for understanding player dynamics. This report will navigate through various predictive models, hypothesis testing, and data visualizations to offer insights into the correlations between player attributes and their professional standing.

**Importing the necessary libraries**

```{r message=FALSE}
library(dplyr)
library(corrplot)
library(tidyr)
list.of.packages <- c( "ggplot2", "pls", "Matrix", "caTools", "glmnet")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
require(glmnet)
require(caTools)
library(tidyverse)
library(GGally)
```

------------------------------------------------------------------------

# II. Methods

|     |                                                                                                                                                                                                                                 |                                              |
|-----|---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|----------------------------------------------|
|     | **Question**                                                                                                                                                                                                                    | **Method**                                   |
| Q1  | Can we predict the Overall Performance Rating of football players based on key performance attributes such as pace, shooting, and dribbling?                                                                                    | Predictive Modeling with Multiple Regression |
| Q2  | Can we predict a player's 'club_position' (e.g., Goalkeeper, Defender, Midfielder, Forward) based on their individual performance attributes such as pace, shooting, passing, dribbling, defending, physic, overall, potential? | Predictive Modeling with Classification      |
| Q3  | Can we predict a player's average pace (a combination of acceleration and sprint speed) based on their age?                                                                                                                     | Hypothesis testing                           |
| Q4  | Can we predict a player's market value based on attributes such as age, overall rating, potential, and specific skill attributes (passing, dribbling, shooting) and other variables?                                            | Predictive Modeling with Regression          |

------------------------------------------------------------------------

# III. Analysis

```{r}
# Loading the dataset
df <- read.csv("players_22.csv")
```

## Business Questions

[**Question 1:**]{.underline} **Can we predict the Overall Performance Rating of football players based on key performance attributes such as pace, shooting, and dribbling?**

-   **Exploratory Data Analysis**

```{r}
#Check for missing values
missing_values <- colSums(is.na(df))
names(missing_values)

#Handle missing values (e.g., fill with mean or median)
df$pace[is.na(df$pace)] <- mean(df$pace, na.rm = TRUE)
df$shooting[is.na(df$shooting)] <- mean(df$shooting, na.rm = TRUE)
df$dribbling[is.na(df$dribbling)] <- mean(df$dribbling, na.rm = TRUE)

#Winsorizing outliers in the 'pace' column
q<-quantile(df$pace, c(0.01, 0.99), na.rm = TRUE)
df$pace[df$pace < q[1]] <- q[1]
df$pace[df$pace > q[2]] <- q[2]

```

```{r}
# Boxplots for individual attributes
df %>%
  pivot_longer(cols = c("pace", "shooting", "dribbling"), names_to = "Attribute", values_to = "Value") %>%
  ggplot(aes(x = Attribute, y = Value)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Figure 1.1: Boxplots for Attributes",
       caption = "Note:The box represents the interquartile range (IQR), with the line inside indicating the median value.",
       x = "Attribute",
       y = "Value") +
  theme_minimal()

```

**Interpretation of Figure 1.1: Boxplots for Attributes**

-   The boxplots show the distribution of values for each attribute (pace, shooting, and dribbling).

-   The box represents the interquartile range (IQR), with the median line inside the box.

-   Whiskers extend to the minimum and maximum values within 1.5 times the IQR from the lower and upper quartiles. Points beyond the whiskers are considered potential outliers.

```{r}
#Correlation
cor_matrix <- cor(df[c("pace", "shooting", "dribbling")])

#Correlation matrix visualization
corrplot(cor_matrix, method = "circle", title= "Figure 1.2: Correlation Matrix",mar = c(0,0,1,0), tl.srt = 45, tl.cex = 0.8, cex.main = 0.9) 

# Add a note below the plot
mtext("Note:Stronger correlations are indicated by larger circle sizes and darker colors.",   side = 1, line = 4, cex = 0.7)
```

**Interpretation of Figure 1.2: Correlation Matrix**

-   The correlation matrix is a table that shows the correlation coefficients between different variables. In this case, it represents the correlation between the attributes: pace, shooting, and dribbling.

-   The values in the matrix range from -1 to 1. A value closer to 1 indicates a strong positive correlation, while a value closer to -1 indicates a strong negative correlation. A value of 0 suggests no linear correlation.

-   The color intensity and the size of the circles in the matrix provide a visual representation of the correlation strength. Larger and darker circles signify stronger correlations.

-   Use the correlation matrix to identify relationships between attributes. For example, a high positive correlation between pace and dribbling suggests that players who are faster may also have better dribbling skills.

------------------------------------------------------------------------

-   **Methodology- Predictive Modeling with Multiple Regression**

    1.  **Data Preprocessing:**

        -   Identify and handle missing values in the key performance attributes (e.g., pace, shooting, dribbling) and the target variable (Overall Performance Rating).

        -   Examine potential outliers in the performance attributes and the target variable. Decide on an appropriate strategy for treating outliers, such as transformation or removal.

        -   If the key performance attributes have different scales, consider scaling or normalizing them to ensure that they contribute equally to the regression model.

    2.  **Feature Selection:**

        -   Assess the correlation between key performance attributes and the target variable. Select attributes with strong correlations for inclusion in the model.

    3.  **Splitting the dataset:**

        -   Divide the dataset into training and testing sets. This allows you to train the model on one subset and evaluate its performance on another, providing an estimate of the model's generalization to new data. Split, considering 70% for training and 30% for testing.

    4.  **Model Building:**

        -   Implement multiple regression by including the selected key performance attributes (e.g., pace, shooting, dribbling) as predictors in the model.

    5.  **Model Evaluation:**

        -   Evaluate the model using appropriate metrics such as Mean Squared Error (MSE) and Root Mean Squared Error (RMSE).

-   **Why use Multiple Regression?**

    Multiple regression allows you to examine how changes in each key performance attribute relate to changes in the Overall Performance Rating while holding other variables constant. This is important when assessing the unique contribution of each attribute. The inclusion of multiple relevant predictors can enhance the model's predictive accuracy. By incorporating a combination of key performance attributes, the model can better capture the variability in the Overall Performance Rating.

------------------------------------------------------------------------

------------------------------------------------------------------------

[**Question 2:**]{.underline} **Can we predict a player's 'club_position' (e.g., Goalkeeper, Defender, Midfielder, Forward) based on their individual performance attributes such as pace, shooting, passing, dribbling, defending, physic, overall, potential?**

-   **Exploratory Data Analysis**

```{r}
skimr::skim(df)
```

```{r}
# Creating a column with simplified positions since our column "club_postions" has more than 10 postions such as LW,RW,ST,CF,CAM,CDM,GK,CB,RB etc. 

df$simplified_position <- dplyr::case_when(
  df$club_position %in% c("GK") ~ "Goalkeeper",
  df$club_position %in% c("LCB", "RCB", "CB", "LB", "RB", "LWB", "RWB", "RES") ~ "Defender",
  df$club_position %in% c("RCM", "LCM", "CDM", "RDM", "LDM", "CM", "CAM", "RM", "LM", "RAM", "LAM") ~ "Midfielder",
  df$club_position %in% c("RW", "ST", "LW", "CF", "RS", "LS", "RF", "LF", "SUB") ~ "Forward",
  TRUE ~ as.character(df$club_position) 
)

# Converting the new column to a factor
df$simplified_position <- as.factor(df$simplified_position)

#Data Cleaning- Removing rows with "" values
matches <- grepl("", df$simplified_position)
df <- df[df$simplified_position != "", ]

# Plotting bargraph 
ggplot(df, aes(x = simplified_position)) +
  geom_bar() +  # This creates the bar plot
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5, position = position_stack(vjust = 0.5)) +  # This adds the count labels
  theme_minimal() +
  labs(title = "Figure 2.1: Distribution of Simplified Club Positions", x = "Simplified Club Position",
       caption = "Note : Bar plot illustrating the distribution of players across simplified club positions within the dataset. ",y = "Count")
```

**Interpretation of Figure 2.1:**

This chart illustrates the count of players in a dataset categorized by four simplified club positions: Defender, Forward, Goalkeeper, and Midfielder. The position of Forward is the most common among this group of players. The data contains a relatively balanced number of Defenders and Midfielders, with Defenders being more numerous. Goalkeepers are the least represented in the dataset, which aligns with the structure of soccer teams where there are usually fewer goalkeepers than players for other positions.

```{r}
# Select only numerical features for the pairwise comparison
features <- c('pace', 'shooting', 'passing', 'dribbling', 'defending', 'physic', 'overall', 'potential')
df_numeric <- df %>%
  select(all_of(features)) %>%
  select_if(is.numeric) %>%
  na.omit()


# For boxplots of each feature by simplified_position, we melt the data and use ggplot
df_melted <- df %>%
  select(simplified_position, all_of(features)) %>%
  gather(key = "feature", value = "value", -simplified_position) %>% # Convert to long format
  na.omit() # Omit missing values

ggplot(df_melted, aes(x = simplified_position, y = value)) +
  geom_boxplot() +
  facet_wrap(~feature, scales = 'free_y') + # Create a separate plot for each feature
  theme_minimal() +
  labs(title = "Figure 2.2: Performance Metrics by Club Position", x = "Simplified Club Position",
       caption = "Note: Box plots displaying the distribution of various performance metrics  across simplified club positions ",y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

**Interpretation of Figure 2.2:**

-   **Defenders** are rated highest in defending, which is consistent with their role in preventing the opposition from scoring.

-   **Forwards** excel in shooting and pace, indicating their role in fast attacks and goal-scoring opportunities.

-   **Goalkeepers** show high ratings in defending, underscoring their role in stopping shots on goal, but they have lower scores in outfield skills like pace, shooting, and dribbling.

-   **Midfielders** have the highest ratings in passing, reflecting their central role in creating plays and distributing the ball.

Overall ratings are relatively even across all positions, suggesting a balanced distribution of general skill levels. The metrics align well with the typical skillsets required for each position in football.

------------------------------------------------------------------------

-   **Methodology- Predictive Modeling with Classification**

    1.  **Data Preprocessing:**

        -   Clean the dataset by handling missing values, outliers, and inconsistencies.

        -   Normalize or scale numerical features to ensure uniformity.

    2.  **Feature Selection:**

        -   Identify and select relevant features (as stated above) that significantly contribute to predict 'club_position.'

    3.  **Splitting the dataset:**

        -   Divide the dataset into training and testing sets to evaluate the model's performance accurately. Split, considering 70% for training and 30% for testing.

    4.  **Model Building:**

        -   Implement the K-Nearest Neighbors (KNN) algorithm for the chosen features.

        -   Train the model using the training dataset.

    5.  **Model Evaluation:**

        -   Evaluate the model's performance on the testing dataset using appropriate metrics such as accuracy, precision, recall, and F1 score.

-   **Why using K-Nearest Neighbors (KNN)?**

    KNN is suitable for classifying players into positions due to its simplicity, effectiveness with complex decision boundaries, and flexibility in handling various data types without assuming data distribution, making it ideal for the diverse and nuanced attributes of football players.

    This classification question is important as it relates to determining a player's most suitable position on the football field based on their skill set and physical attributes. The ability to accurately classify a player's position can be invaluable for team formation, scouting, and tactical analysis.

------------------------------------------------------------------------

------------------------------------------------------------------------

[**Question 3:**]{.underline} **Can we predict a player's average pace (a combination of acceleration and sprint speed) based on their age?**

-   **Exploratory Data Analysis**

    ```{r}
    # Adding column indicating combination of acceleration and speed
    df<- df%>%
      mutate(average_pace = (movement_acceleration + movement_sprint_speed) / 2)

    # Calculate median
    median_age <- median(df$age)

    # Plotting histogram showing the age distribution
    ggplot(df, aes(x = age)) +
      geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
      geom_vline(xintercept = median_age, color = "red", linetype = "dashed", linewidth = 0.5) + 
      labs(title = paste("Figure 3.1: Distribution of Age with Median (", median_age, ") Line"), 
           x = "Age", y = "Frequency") +
      labs(caption = "Note: The dashed red line indicates the median age, providing insight into the central tendency of the distribution.")


    ```

**Interpretation of Figure 3.1:**

The histogram shows that the ages in the dataset are spread out across a range from approximately 15 to 50 years old. The distribution of age is not symmetrical; it is left-skewed, indicating that a larger portion of the dataset is composed of younger individuals. The dashed red line represents the median age, which is 25 years, hence we will consider 25 years for the hypothesis. The skewness of the distribution suggests that the population has a younger demographic, with fewer individuals in the older age groups.

```{r}
# Plotting the boxplot of average pace by age group
ggplot(df, aes(x = factor(cut(age, breaks = c(0, 25, Inf))), y = average_pace, fill = factor(cut(age, breaks = c(0, 25, Inf))))) +
  geom_boxplot() +
  labs(title = "Figure 3.2: Boxplot of Average Pace by Age Group", x = "Age Group", y = "Average Pace", fill = "Age Group",
       caption = "Note: This boxplot visualizes the distribution of average pace across different age groups.")

```

**Interpretation of Figure 3.2:**

The first group includes ages from just above 0 up to and including 25, while the second group includes ages above 25 up to infinity (or the maximum age in the dataset). Both boxes seem to have a similar range of average pace, indicated by the height of the boxes, which represents the interquartile range (IQR), the middle 50% of the data. The median pace (indicated by the line within each box) appears to be slightly higher for the older age group than for the younger one, suggesting that the median individual over 25 years old has a slower pace.

```{r}
# Plotting cumulative distribution
ggplot(df, aes(x = average_pace, color = factor(ifelse(age < 25, "Under 25", "25 and Above")))) +
  stat_ecdf(geom = "step", size = 0.75) +
  labs(title = "Figure 3.3: Cumulative Distribution Function of Average Pace by Age Group", color = "Age Group") +
  labs(caption = "Note:The colored lines represent age groups, distinguishing players under 25 from those 25 and above.")

```

**Interpretation of Figure 3.3:**

The lines for the two groups are close together, indicating that the distributions of average pace for both age groups are similar. At many points across the distribution, the line for the "25 and Above" group is slightly higher than the line for the "Under 25" group. This suggests that in general, a higher proportion of the "25 and Above" group have slower paces compared to the "Under 25" group up until the faster paces are reached where the lines converge.

------------------------------------------------------------------------

-   **Methodology- Hypothesis Testing**

    1.  **Data Preprocessing:**

        -   Clean the dataset by handling missing values, outliers, and inconsistencies.

    2.  **Hypothesis Formulation:**

        -   Formulate the null hypothesis (H0) that there is no significant difference in average pace between players under 25 and those 25 years old or above.

        -   Formulate the alternative hypothesis (H1) that there is a significant difference in average pace based on age groups.

    3.  **Splitting the dataset:**

        -   Split the dataset into two groups: players under 25 years old and players 25 years old or above.

    4.  **Hypothesis Testing:**

        -   Perform a t-test or an appropriate statistical test to compare the average pace between the two age groups.

        -   Determine the p-value and assess its significance level (commonly set at 0.05) to decide whether to reject the null hypothesis.

    5.  **Results Interpretation:**

        -   Interpret the results of hypothesis testing, stating whether there is a significant difference in average pace between the age groups.

-   **Why using Hypothesis Testing?**

    Hypothesis testing is used in the methodology outlined above to scientifically assess whether there is a significant difference in average pace between two specific age groups of football players (those under 25 and those 25 or above). By conducting hypothesis testing, we can determine whether any observed differences in average pace between the age groups are statistically significant. This allows us to make informed conclusions about the impact of age on average pace.

------------------------------------------------------------------------

------------------------------------------------------------------------

[**Question 4:**]{.underline} **Can we predict a player's market value based on attributes such as age, overall rating, potential, and specific skill attributes (passing, dribbling, shooting) and other variables?**

-   **Exploratory Data Analysis**

```{r}

# Impute missing values with mean for each column (numeric columns only)
numeric_df <- df[sapply(df, is.numeric)]

numeric_df_imputed <- numeric_df

# Apply the mean for each column
for(i in seq_along(numeric_df_imputed)) {
  numeric_df_imputed[[i]] <- ifelse(is.na(numeric_df_imputed[[i]]), 
                                    mean(numeric_df_imputed[[i]], na.rm = TRUE), 
                                    numeric_df_imputed[[i]])
}

# Calculate the correlation matrix
cor_matrix <- cor(numeric_df_imputed, use = "complete.obs")

```

```{r}

market_value_correlations <- cor_matrix[,'value_eur']
sorted_correlations <- sort(market_value_correlations, decreasing = TRUE)
print(sorted_correlations)

# Filtering high correlations, as an example
high_correlations <- sorted_correlations[sorted_correlations > 0.5 | sorted_correlations < -0.5]

# Printing top 5
names(high_correlations)
```

```{r}

high_corr_vars <- names(high_correlations)

# Subset the correlation matrix to include only the highly correlated variables
high_corr_matrix <- cor_matrix[high_corr_vars, high_corr_vars]

# Generate the corrplot
corrplot::corrplot(high_corr_matrix, method = "color", type = "upper", 
                   tl.col = "black", tl.srt = 30, diag = FALSE,
                   title = "Figure 4.1 - Correlation Matrix of Highly Correlated Variables with Market Value",
                   mar = c(0,0,1,0), tl.cex = 0.8, cex.main = 0.9) 

# Add a note below the plot
mtext("Note: Shows a color-coded correlation matrix of variables that are highly correlated with market value.", 
      side = 1, line = 4, cex = 0.8)
```

**Interpretation of Figure 4.1:**

1.  **Variable Relationships:**

    -   Specific pairs of variables seem to have a very strong positive correlation, as indicated by the dark blue squares. For example, **`value_eur`** is strongly correlated with **`release_clause_eur`**, which suggests that as the market value of a player increases, their release clause value also tends to increase.

    -   The variable **`wage_eur`** also shows a strong positive correlation with **`value_eur`**, indicating that players with higher market values tend to have higher wages.

    -   **`International_reputation`** and **`overall`** show a strong positive correlation with **`value_eur`** as well, which implies that players with higher overall ratings and international reputation are likely to have higher market values.

2.  **Potential for Multicollinearity:**

    -   There may be multicollinearity present, as indicated by strong correlations between independent variables (e.g., **`overall`** and **`international_reputation`**). This multicollinearity can be a problem for regression models as it can inflate the variance of coefficient estimates and make the model unstable.

3.  **Interpretation for Regression:**

    -   For predictive modeling, variables with high correlation to **`value_eur`** are prime candidates for inclusion in a regression model. However, care must be taken to assess the impact of multicollinearity and to interpret the coefficients correctly.

```{r}

# Transform the 'value_eur' with a log transformation to handle skewness
df$value_eur_log <- log1p(df$value_eur)  # log1p is used to handle zero values

# Create the histogram with log-transformed data
ggplot(df, aes(x = value_eur_log)) +
  geom_histogram(binwidth = 0.1, fill = "blue", alpha = 0.7) +
  scale_x_continuous(name = "Market Value (EUR)", 
                     breaks = scales::pretty_breaks(n = 10), 
                     labels = scales::comma) +
  scale_y_continuous(name = "Frequency", 
                     labels = scales::comma) +
  labs(title = "Figure 4.2 - Histogram of Distribution of Players' Market Value", caption = "Note: The above Histogram shows the distribution of Player's Market Value after a log transformation ") +
  theme_minimal() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))
```

**Interpretation of Figure 4.2:**

The histogram depicts the distribution of players' market values after a log transformation, which is evident from the x-axis being labeled with log-scaled values. The distribution shows that the majority of players have market values concentrated in the middle range of the log scale, with fewer players as the market value increases. The shape of the distribution appears to be roughly normal with a slight right skew, indicating that high-value players are rarer. The histogram provides a visual representation of how players' market values are spread out, highlighting the commonality of certain value ranges and the relative scarcity of very high values.

```{r}
# Create scatterplot for 'value_eur' vs 'overall'
ggplot(df, aes(x = overall, y = value_eur)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + 
  labs(title = "Figure 4.3 - Scatterplot of Overall Rating vs Market Value",
       caption= "Note: The above scatterplot shows relationship between 'Overall Rating' and 'Market Value (EUR) ",
       x = "Overall Rating",
       y = "Market Value (EUR)")


```

**Interpretation of Figure 4.3:**

This scatterplot demonstrates the relationship between football players' overall ratings and their market values. Each dot represents an individual player, plotted with their overall rating on the x-axis and their market value on the y-axis. There is a trend line indicating a positive correlation; as the overall rating increases, the market value tends to increase as well. The spread of points widens at higher ratings, suggesting greater variance in market value among top-rated players.

```{r}

# Create scatterplot for 'value_eur' vs 'international_reputation'
ggplot(df, aes(x = international_reputation, y = value_eur)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + 
  labs(title = "Figure 4.4 - Scatterplot of International Reputation vs Market Value",caption= "Note: The above scatterplot shows relationship between 'International Reputation' and 'Market Value (EUR) ",
       x = "International Reputation",
       y = "Market Value (EUR)")
```

**Interpretation of Figure 4.4:**

The scatterplot illustrates the relationship between football players' international reputation and their market values. Each point represents a player, with international reputation on the x-axis and market value on the y-axis. There appears to be a positive correlation: as international reputation increases, market value tends to rise. The trend line reinforces this positive relationship. It's also noticeable that players with the highest international reputation have a wider range of market values, which may suggest that factors other than reputation contribute significantly to the market value at the highest levels of reputation.

```{r}
# Scatterplot for 'release_clause_eur' vs 'value_eur'
ggplot(df, aes(x = release_clause_eur, y = value_eur)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + 
  labs(title = "Figure 4.5 - Scatterplot of Release clause vs Market Value",caption= "Note: The above scatterplot shows relationship between 'Release Clause' and 'Market Value (EUR) ",
       x = "Release Clause",
       y = "Market Value (EUR)")
```

**Interpretation of Figure 4.5:**

The scatterplot presents a comparison between football players' release clause amounts and their market values. The trend is strongly positive, showing that players with higher release clauses generally have higher market values. The tight clustering of data points along the trend line indicates a strong linear relationship, suggesting that the release clause is a good predictor of market value. This could imply that the release clause is set in close consideration of a player's perceived market value.

```{r}

# Scatterplot for 'wage_eur' vs 'value_eur'
ggplot(df, aes(x = wage_eur, y = value_eur)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + 
  labs(title = "Figure 4.6 - Scatterplot of Wage vs Market Value",caption= "Note: The above scatterplot shows relationship between 'Wage' and 'Market Value (EUR) ",
       x = "Wage",
       y = "Market Value (EUR)")
```

**Interpretation of Figure 4.6:**

The scatterplot displays the relationship between football players' wages and their market values. The data points suggest a positive correlation, as players with higher wages also tend to have higher market values. The trend line indicates this positive trajectory, although the spread of data points suggests that while wage is a significant predictor of market value, there are other factors also influencing a player's value, given the variation at different wage levels.

```{r}
# Scatterplot for 'potential' vs 'value_eur'
ggplot(df, aes(x = potential, y = value_eur)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + 
  labs(title = "Figure 4.7 - Scatterplot of Potential vs Market Value",caption= "Note: The above scatterplot shows relationship between 'Potential' and 'Market Value (EUR) ",
       x = "Potential",
       y = "Market Value (EUR)")
```

**Interpretation of Figure 4.7:**

The scatterplot demonstrates the relationship between the potential of football players and their market values. It shows that players with higher potential tend to have higher market values. The distribution of points indicates a positive correlation, particularly noticeable at the higher end of the potential scale where the market values vary more widely. This suggests that while potential is a strong indicator of market value, other factors also contribute to a player's valuation, especially among those with the highest potential.

------------------------------------------------------------------------

-   **Methodology- Predictive Modeling with Regression**

    1.  **Data Preprocessing:**

        -   Clean the dataset by handling missing values, outliers, and inconsistencies.

        -   Scale numerical features to ensure that attributes like age, overall rating, potential, and skill attributes have equal weight.

    2.  **Feature Selection:**

        -   Identify and select relevant features that significantly contribute to determining a player's market value.

    3.  **Splitting the dataset:**

        -   Divide the dataset into training and testing sets to evaluate the model's performance accurately. An 70:30 split is commonly used, allocating 70% for training and 30% for testing

    4.  **Model Building:**

        -   Implement Ridge and Lasso regression models for the chosen features, tuning the regularization strength via hyperparameter tuning to prevent overfitting and to handle multicollinearity.

        -   Train the models using the training dataset, ensuring that the regularization parameter is appropriately calibrated using techniques such as cross-validation.

    5.  **Model Evaluation:**

        -   Evaluate the model's performance on the testing dataset using appropriate metrics such as R-squared, Mean Absolute Error (MAE), Mean Squared Error (MSE), and Root Mean Squared Error (RMSE).

-   **Why using Ridge and Lasso Regression?**

Ridge and Lasso regression are used due to their ability to handle multicollinearity between independent variables, which is common in datasets with many features. They help in feature selection (Lasso) and in shrinking the coefficients (Ridge) to minimize overfitting, leading to a more generalizable model. They are particularly useful when there is a suspicion that not all features are relevant or when we have more features than observations, which can be the case in high-dimensional datasets often found in player performance data. Lasso Regression can zero out the less important features' coefficients, effectively performing feature selection, which simplifies the model and makes it easier to interpret. Ridge Regression is used when we want to include all features but penalize their coefficients, which helps in dealing with overfitting and improving model performance.

------------------------------------------------------------------------

# IV. Conclusion

In conclusion, the analytical journey through the FIFA 2022 dataset has provided us with a nuanced understanding of the factors influencing a player's market value and performance. Through multiple regression models and classification methods, we've been able to predict overall performance ratings and club positions with significant accuracy. Furthermore, hypothesis testing has shed light on the influence of age on a player's pace. The strong correlation between players' market values and their on-field attributes such as potential, overall rating, and international reputation underscores the multi-faceted nature of player valuation. This analysis serves as a testament to the power of data-driven approaches in sports analytics, offering valuable insights that can inform decision-making in football management and player investment.

------------------------------------------------------------------------

# V. References

-   A host of comprehensive sports datasets for research, analysis, data modeling, data-visualization, predictions, machine-learning ETC. Sports. (n.d.). <https://sports-statistics.com/sports-data/sports-data-sets-for-data-modeling-visualization-predictions-machine-learning/> 

-   Developing research questions. Library. (2023, March 3). <https://www.monash.edu/library/help/assignments-research/developing-research-questions> 
